---
title: "Data Prep"
author: "Leonor Brito"
date: "2024-02-20"
output: pdf_document
---

```{r setup}

knitr::opts_knit$set(root.dir = "/Users/leonorbrito/Documents/Master/Thesis/Datasets/macro")

```

```{r}

library(dplyr)
library(tidyr)
library(purrr)

```

# Import dataframes
```{r}

crop_nutrient_balance <- read.csv('Environment_Cropland_nutrient_budget_E_All_Data.csv')
land_use <- read.csv('Inputs_LandUse_E_All_Data.csv')
land_cover <- read.csv('Environment_LandCover_E_All_Data.csv')
fertilizer_use <- read.csv('Inputs_FertilizersNutrient_E_All_Data.csv')
pesticide_use <- read.csv('Inputs_Pesticides_Use_E_All_Data.csv')
temp_change <- read.csv('Environment_Temperature_change_E_All_Data.csv', fileEncoding = "ISO-8859-1")
employment <- read.csv('Employment_Indicators_Agriculture_E_All_Data.csv')
fertilizer_product <- read.csv('Inputs_FertilizersProduct_E_All_Data.csv')
manure <- read.csv('Environment_LivestockManure_E_All_Data.csv')

```

# Data Clean and Prep

## Remove unwanted year columns from all datasets
```{r}

remove_year_FN_columns <- function(df) {
  # Identify columns that end with 'F' or 'N' following a year (assumes year is four digits)
  pattern <- "Y[0-9]{4}[FN]$"
  columns_to_remove <- grep(pattern, names(df), value = TRUE)
  
  # Remove the identified columns from the dataframe
  df <- df[, !names(df) %in% columns_to_remove]
}

land_cover <- remove_year_FN_columns(land_cover)
land_use <- remove_year_FN_columns(land_use)
employment <- remove_year_FN_columns(employment)
pesticide_use <- remove_year_FN_columns(pesticide_use)
temp_change <- remove_year_FN_columns(temp_change)
crop_nutrient_balance <- remove_year_FN_columns(crop_nutrient_balance)
manure <- remove_year_FN_columns(manure)
fertilizer_product <- remove_year_FN_columns(fertilizer_product)
fertilizer_use <- remove_year_FN_columns(fertilizer_use)

```

## Function to check for NAs
```{r}

calculate_na_sums <- function(dataframe, variable_column) {
  # Identify year columns based on a naming convention (e.g., starting with 'Y_')
  year_columns <- grep("^Y", names(dataframe), value = TRUE)
  
  # Function to calculate sum of NAs for a given variable value within the specified year columns
  sum_nas_by_variable <- function(variable_value) {
    subset_df <- subset(dataframe, dataframe[[variable_column]] == variable_value)
    # Use lapply to handle NA counting, then unlist to ensure the result is a numeric vector
    na_counts <- unlist(lapply(subset_df[year_columns], function(x) sum(is.na(x))))
    sum_na <- sum(na_counts)
    return(sum_na)
  }
  
  # Get unique variable values from the specified variable column
  variables <- unique(dataframe[[variable_column]])
  
  # Calculate the sum of NAs for each variable value
  na_sums <- sapply(variables, sum_nas_by_variable)
  
  # Name the result vector for clarity
  names(na_sums) <- variables
  
  return(na_sums)
}

```

### Land Cover

```{r}

calculate_na_sums(land_cover, "Element")

```

```{r}

# Identify rows for "Area from CCI_LC" with NAs in any of the year columns
cci_lc_rows <- land_cover$Element == "Area from CCI_LC"
year_columns <- grep("^Y", names(land_cover), value = TRUE)
na_indices_cci_lc <- which(cci_lc_rows & rowSums(is.na(land_cover[year_columns])) > 0)

# Initialize a list to store results
data_points_in_others <- list()

# Loop through each element value excluding "Area from CCI_LC"
unique_elements <- unique(land_cover$Element[land_cover$Element != "Area from CCI_LC"])
for(elem in unique_elements) {
  # For each element, check corresponding rows in year columns for non-NA where CCI_LC has NA
  elem_rows <- land_cover$Element == elem
  # Subset land_cover for this element and year columns, then check for NAs
  elem_data <- land_cover[elem_rows, year_columns]
  
  # Count how many of the NA indices for CCI_LC have non-NA values for this element
  non_na_counts <- sapply(na_indices_cci_lc, function(index) {
    sum(!is.na(elem_data[index, ]))
  })
  
  # Sum to get total non-NA counts for this element where CCI_LC has NA
  total_non_na_for_elem <- sum(non_na_counts)
  
  # Store in the list with the element name as the key
  data_points_in_others[[elem]] <- total_non_na_for_elem
}

# Print the list to see the counts
print(data_points_in_others)


```

```{r}

# Step 1: Convert the dataset to long format including the Year columns
land_cover_long <- land_cover %>%
  pivot_longer(cols = starts_with("Y"), 
               names_to = "Year", 
               values_to = "Value",
               names_prefix = "Y") %>%
  mutate(Year = as.numeric(Year)) # Ensure Year is numeric

# Step 2: Group by Area, Item, and Year to calculate the average across Elements
# Note that we assume 'Element' is also a column in your dataset
land_cover_avg <- land_cover_long %>%
  group_by(Area.Code, Area, Item, Year) %>%
  summarise(Average_Value = mean(Value, na.rm = TRUE), .groups = 'drop')

# Step 3: Spread the Item column to wide format
land_cover_final <- land_cover_avg %>%
  pivot_wider(names_from = Item, 
              values_from = Average_Value)

```

### Land Use

```{r}

# Define the combinations of interest
combinations_of_interest <- data.frame(
  Item = c("Agricultural land", "Cropland", "Arable land", "Permanent crops", 
           "Land area equipped for irrigation", "Land area actually irrigated", 
           "Agriculture area under organic agric.", "Forest land"),
  Element = c("Share in Land area", "Share in Agricultural land", "Share in Agricultural land", 
              "Share in Agricultural land", "Share in Cropland", 
              "Area", 
              "Share in Agricultural land", "Share in Land area")
)

# Filter the land_use dataframe to keep only the rows that match the combinations of interest
filtered_land_use <- land_use %>%
  semi_join(combinations_of_interest, by = c("Item", "Element"))

# Get unique countries from the land_use data
unique_countries <- unique(land_use$Area)

# For each item-element combination, calculate how many countries it's missing from
missing_combinations_count <- combinations_of_interest %>%
  rowwise() %>%
  mutate(MissingInCountriesCount = length(unique_countries) - 
         sum(land_use$Item == Item & land_use$Element == Element)) %>%
  ungroup()

# View the result
print(missing_combinations_count)

```
Remove "Land area actually irrigated"
```{r}

filtered_land_use <- filtered_land_use %>%
  filter(Item != "Land area actually irrigated")

```
Concatenate Element and Item to make one variable column
```{r}

filtered_land_use <- filtered_land_use %>%
  mutate(combined = paste(Item, Element, sep = ", "))

# Remove unwanted columns
filtered_land_use <- filtered_land_use %>%
  select(-c(Item.Code, Item, Element.Code, Element, Area.Code..M49., Unit))

```
Convert variable to wide format, years to long format
```{r}

# Pivot years columns to long format
land_use_long <- filtered_land_use %>%
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "Year",
    values_to = "Value"
  )

# Pivot the 'combined' column to wide format
land_use_final <- land_use_long %>%
  pivot_wider(
    names_from = combined,
    values_from = Value
  )

```

### Pesticide Use

```{r}

pesticide_use_filtered <- pesticide_use %>%
  filter(Item == "Pesticides (total)" & Element == "Use per area of cropland")

# concatenate Item and Element to create a new variable
pesticide_use_filtered <- pesticide_use_filtered %>%
  mutate(combined = paste(Item, Element, sep = ", "))

# remove unwanted columns
pesticide_use_filtered <- pesticide_use_filtered %>%
  select(-c(Item.Code, Item, Element.Code, Element, Area.Code..M49., Unit))

```

Convert variable to wide format, years to long format
```{r}

# Pivot years columns to long format
pesticide_use_long <- pesticide_use_filtered %>%
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "Year",
    values_to = "Value"
  )

# Pivot the 'combined' column to wide format
pesticide_use_final <- pesticide_use_long %>%
  pivot_wider(
    names_from = combined,
    values_from = Value
  )

```

### Temp Change

```{r}

temp_change_filtered <- temp_change %>%
  filter(Months == "Meteorological year", Element == "Temperature change")

```

Convert variable to wide format, years to long format
```{r}

# Pivot years columns to long format
temp_change_long <- temp_change_filtered %>%
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "Year",
    values_to = "Value"
  )

# Pivot the 'combined' column to wide format
temp_change_final <- temp_change_long %>%
  pivot_wider(
    names_from = Element,
    values_from = Value
  )


names(temp_change_final)[names(temp_change_final) == "Temperature change"] <- "Temperature change (C)"

# remove unwanted columns
temp_change_final <- temp_change_final %>%
  select(-c(Area.Code..M49., Months.Code, Months, Element.Code, Unit))

```


### Crop Nutrient Balance

```{r}

# Filter the crop_nutrient_balance dataset for nue Element values
filtered_nue <- subset(crop_nutrient_balance, Element %in% c("Cropland nitrogen use efficiency", 
                                                            "Cropland phosphorus use efficiency", 
                                                            "Cropland potassium use efficiency"))

# Pivot years columns to long format
nue_long <- filtered_nue %>%
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "Year",
    values_to = "Value"
  )

# Pivot the 'combined' column to wide format
nue_final <- nue_long %>%
  pivot_wider(
    id_cols = c(Area.Code, Area, Year),
    names_from = Element,
    values_from = Value
  )


```

### Manure Management

```{r}

# Filter the crop_nutrient_balance dataset for nue Element values
filtered_manure <- subset(manure, Element %in% c("Stocks", "Amount excreted in manure (N content)", "Manure left on pasture (N content)"))

# Pivot years columns to long format
manure_long <- filtered_manure %>%
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "Year",
    values_to = "Value"
  )

# Pivot the 'combined' column to wide format
manure_wide <- manure_long %>%
  pivot_wider(
    id_cols = c(Area.Code, Area, Item, Year),
    names_from = Element,
    values_from = Value
  )


# Calculate averages and remove the "Item" column
manure_final <- manure_wide %>%
  group_by(Area.Code, Area, Year) %>%
  summarise(
    Avg_Stocks = mean(`Stocks`, na.rm = TRUE),
    Avg_Amount_Excreted = sum(`Amount excreted in manure (N content)`, na.rm = TRUE),
    Avg_Manure_Left_On_Pasture = sum(`Manure left on pasture (N content)`, na.rm = TRUE)
  ) %>%
  ungroup()  # Optional, to remove the grouping structure


```

### Fertilizer Management

```{r}

# Filter the crop_nutrient_balance dataset for nue Element values
filtered_fertilizer <- subset(fertilizer_use, Element %in% c("Production", "Import Quantity", "Export Quantity"))

filtered_fertilizer$combined <- paste(filtered_fertilizer$Item, filtered_fertilizer$Element, sep = ", ")


# Pivot years columns to long format
fertilizer_long <- filtered_fertilizer %>%
  pivot_longer(
    cols = starts_with("Y"),
    names_to = "Year",
    values_to = "Value"
  )

# Pivot the 'combined' column to wide format
fertilizer_final <- fertilizer_long %>%
  pivot_wider(
    id_cols = c(Area.Code, Area, Year),
    names_from = combined,
    values_from = Value
  )

# Create import/export ratio variable
fertilizer_final$N.ImportExport.Ratio <- fertilizer_final[["Nutrient nitrogen N (total), Import Quantity"]] / fertilizer_final[["Nutrient nitrogen N (total), Export Quantity" ]]
fertilizer_final$P.ImportExport.Ratio <- fertilizer_final[["Nutrient phosphate P2O5 (total), Import Quantity"]] / fertilizer_final[["Nutrient phosphate P2O5 (total), Export Quantity"]]
fertilizer_final$K.ImportExport.Ratio <- fertilizer_final[["Nutrient potash K2O (total), Import Quantity"]] / fertilizer_final[["Nutrient potash K2O (total), Export Quantity"]]

fertilizer_final <- fertilizer_final %>%
  select(-c(5, 6, 8, 9, 11, 12))

```


# Combine all datasets

```{r}

# Function to convert Year column: remove "Y" prefix and convert to numeric
convert_year <- function(df) {
  df %>%
    mutate(Year = as.numeric(sub("Y", "", Year))) %>%
    return()
}

# Apply the function to each dataset
fertilizer_final <- convert_year(fertilizer_final)
land_cover_final <- convert_year(land_cover_final)
land_use_final <- convert_year(land_use_final)
manure_final <- convert_year(manure_final)
nue_final <- convert_year(nue_final)
pesticide_use_final <- convert_year(pesticide_use_final)
temp_change_final <- convert_year(temp_change_final)


# Now, perform the full joins
combined_final <- fertilizer_final %>%
  full_join(land_cover_final, by = c("Area.Code", "Area", "Year")) %>%
  full_join(land_use_final, by = c("Area.Code", "Area", "Year")) %>%
  full_join(manure_final, by = c("Area.Code", "Area", "Year")) %>%
  full_join(nue_final, by = c("Area.Code", "Area", "Year")) %>%
  full_join(pesticide_use_final, by = c("Area.Code", "Area", "Year")) %>%
  full_join(temp_change_final, by = c("Area.Code", "Area", "Year"))


# Standardize variable column names
year_col_pos <- which(names(combined_final) == "Year")
names(combined_final)[(year_col_pos+1):ncol(combined_final)] <- variable_names

```


```{r}

# Define the variable names
variable_names <- c(
  "n.fert.prod", "p.fert.prod", "k.fert.prod", 
  "n.fert.ratio", "p.fert.ratio", "k.fert.ratio",
  "artif.surf", "coast.water", "grassland", 
  "herb.crops", "inl.water", "magroves", 
  "mult.crops", "snow.areas", "shrub.areas", 
  "flood.areas", "veg.areas", "barren.land", 
  "tree.cover", "woordy.crops", "ag.land.la", 
  "cropland.al", "arable.land.al", "perm.crops.al", 
  "forest.land.al", "irrig.land.cl", "org.land.al", 
  "manure.stocks", "manure.exc", "manure.left", 
  "nue", "pue", "kue", "pest.cl", "temp.change"
)

# Define the descriptions
descriptions <- c(
  "Nutrient nitrogen N (total), Production", "Nutrient phosphate P2O5 (total), Production", "Nutrient potash K2O (total), Production",
  "Nutrient nitrogen N, Import/Export Ratio", "Nutrient phosphate P2O5 (total), Import/Export Ratio", "Nutrient potash K2O (total), Import/Export Ratio",
  "Artificial surfaces (including urban and associated areas)", "Coastal water bodies and intertidal areas", "Grassland",
  "Herbaceous crops", "Inland water bodies", "Mangroves",
  "Multiple or layered crops", "Permanent snow and glaciers", "Shrub-covered areas",
  "Shrubs and/or herbaceous vegetation, aquatic or regularly flooded", "Sparsely natural vegetated areas", "Terrestrial barren land",
  "Tree-covered areas", "Woody crops", "Agricultural land, Share in Land area",
  "Cropland, Share in Agricultural land", "Arable land, Share in Agricultural land", "Permanent crops, Share in Agricultural land",
  "Forest land, Share in Land area", "Land area equipped for irrigation, Share in Cropland", "Agriculture area under organic agric., Share in Agricultural land",
  "Total Manure Stocks", "Amount excreeted in manure (N content)", "Manure left on pasture (N content)",
  "Cropland nitrogen use efficiency", "Cropland phosphorus use efficiency", "Cropland potassium use efficiency",
  "Pesticides (total), Use per area of cropland", "Temperature change"
)

# Define the units
units <- c(
  "t", "t", "t", "t/t", "t/t", "t/t", 
  "1000 ha", "1000 ha", "1000 ha", 
  "1000 ha", "1000 ha", "1000 ha", 
  "1000 ha", "1000 ha", "1000 ha", 
  "1000 ha", "1000 ha", "1000 ha", 
  "1000 ha", "1000 ha", "%", 
  "%", "%", "%", 
  "%", "%", "%", 
  "An", "kg", "kg", 
  "%", "%", "%", 
  "kg/ha", "°c"
)

# Create the data frame
variables_table <- data.frame(Variable = variable_names, Description = descriptions, Unit = units)

# Print the data frame
print(variables_table)


```

```{r}

# List of areas to remove
areas_to_remove <- c(
    "World", "Africa", "Eastern Africa", "Middle Africa", "Northern Africa", "Southern Africa", "Western Africa",
    "Americas", "Northern America", "Central America", "Caribbean", "South America",
    "Asia", "Central Asia", "Eastern Asia", "Southern Asia", "South-eastern Asia", "Western Asia",
    "Europe", "Eastern Europe", "Northern Europe", "Southern Europe", "Western Europe",
    "Oceania", "Australia and New Zealand", "Melanesia", "Micronesia", "Polynesia",
    "European Union (27)", "Least Developed Countries", "Land Locked Developing Countries",
    "Small Island Developing States", "Low Income Food Deficit Countries", "Net Food Importing Developing Countries"
)

# Remove rows where the 'Area' is in the list of areas to remove and Year 2022
combined_final <- combined_final[!combined_final$Area %in% areas_to_remove & combined_final$Year != 2022, ]

```

